Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-02-18 Paper.Rmd
response added

We used datasets that pertain to multiple fields and levels of organization within 
or related to ecology, including environmental, species-trait, phenological, 
population, community, and ecosystem functioning data. Copula structure describing 
the relationship between atmospheric weather variables such as X and Y has been 
examined previously in the meteorological literature (e.g., REF). We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database (RaCA; @Wills2014).
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Measurments were from the central pedon of a soil core, to 1 m depth.
<!--***DAN: Terry to check the wording and accuracy here-->
Species-trait data were average species basal metabolic rate (KJ per hour) and 
body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
Survey suction-trap dataset [@Macaulay1988, @Harrington2014]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years; these are 
indicators of flying aphid abundance. Data pre-processing for the aphid data was the same as 
that of @Sheppard2016,
but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 
years data.  Our second population dataset comprised average annual abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years from 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir 
Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological 
Association of the UK. Data pre-processing was similar but not identical to that 
of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
Community-level data were plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` 
plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 
the year 2000 [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016]. Each included location was measured on at least 50 dates. 
Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
See Appendix \ref{SM-Data_details} for additional data details.

@@

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing the potential for rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewinton-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of populations, the $i$th component $P_i(t)$ representing
the population in the $i$th habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i$th diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is an intrinsic growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed between-patch
Spearman correlations X and Pearson correlations Y in all simulations, 
<!--***Shya: Please fill in X and Y. If it was the Kendall that was used, please 
change Spearman to Kendall. If anything is unclear here in what you should do, please ask. 
If you don't know how to answer these questions, let me know and remind me where the 
code is that you used to generate the noise. Actually, even if you can do this yourself,
please insert a comment below this one indcating where the code is for model and noise-->
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with populations dispersing either to nearest neighbors only,
or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
rate was $d$. After each time step, if the population in a patch was below $1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
among two nearest neighbors (Eq. \ref{eq.extrisk_LC}, $r=0$, $d=$ degree of dispersal). 
$(A)$ Left-tail dependent environmental noise showed higher 
extinction risk than $(B)$ right-tail dependent one, where both case showed extinction risk decreased with 
increasing number of patches from $numlocs =5$ to $25$.\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}
<!--***DAN: Shya, can you please change this figure as follows:
1) Make it only go to d=0.5. I think reviewers will complain that d close to 1 is a weird and 
unrealistic. Anyway, results are the same for larger d, so we are not hiding anything
2) the main comparison is between right and left tail dependence, so make the left panel show 
the numlocs=5 case, with results for both right- and left-tail dependence of the noise shown. 
And make the right panel show the numlocs=25 case with results for both right- and left-tail
dependence of the noise shown.
3) Right now you use grey dots on one panel and black on the other, but you can just use black 
on both.
4)  Use N instead of numlocs - in papers we use variable names like in math, not like in code.
-->

>>>>

We used datasets that pertain to multiple fields and levels of organization within 
or related to ecology, including environmental, species-trait, phenological, 
population, community, and ecosystem functioning data. Copula structure describing 
the relationship between atmospheric weather variables such as X and Y has been 
examined previously in the meteorological literature (e.g., REF). We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database (RaCA; [@Wills2014]).
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Measurments were from the central pedon of a soil core, to 1 m depth.
<!--***DAN: Terry to check the wording and accuracy here-->
Species-trait data were average species basal metabolic rate (KJ per hour) and 
body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
Survey suction-trap dataset [@Macaulay1988; @Harrington2014]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years; these are 
indicators of flying aphid abundance. Data pre-processing for the aphid data was the same as 
that of @Sheppard2016,
but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 
years data.  Our second population dataset comprised average annual abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years from 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir 
Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological 
Association of the UK. Data pre-processing was similar but not identical to that 
of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
Community-level data were plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` 
plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 
the year 2000 [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016]. Each included location was measured on at least 50 dates. 
Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
See Appendix \ref{SM-Data_details} for additional data details.

@@

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing the potential for rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewinton-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of populations, the $i^{th}$ component $P_i(t)$ representing
the population in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is an intrinsic growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed between-patch
Spearman correlations X and Pearson correlations Y in all simulations, 
<!--***DAN: Please fill in X and Y. If it was the Kendall that was used, please 
change Spearman to Kendall. If anything is unclear here in what you should do, please ask. 
If you don't know how to answer these questions, let me know and remind me where the 
code is that you used to generate the noise. Actually, even if you can do this yourself,
please insert a comment below this one indcating where the code is for model and noise
***Shya: I don't know the answer, the code which generates extreme right or left tail dep noise 
is written by you : ExtremeTailDep.R; there is no mention to preserve spearman or kendall or pearson.
I called retd function from ExtremeTaildep.R in another script mydispersal.R line 116, 128 
to generate right and left tail dep. noise.-->
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with populations dispersing either to nearest neighbors only,
or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
rate was $d$. After each time step, if the population in a patch was below $1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
among two nearest neighbors ($r=0$, $d=$ rate of dispersal). 
Left-tail dependent environmental noise caused higher 
extinction risk than right-tail dependent one for number of patches $(A)$ $N = 5$ and $(B)$ $N = 25$, respectively,
where both case showed lower extinction risk with increasing number of patches.\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}
<!--***DAN: Shya, can you please change this figure as follows:
1) Make it only go to d=0.5. I think reviewers will complain that d close to 1 is a weird and 
unrealistic. Anyway, results are the same for larger d, so we are not hiding anything
2) the main comparison is between right and left tail dependence, so make the left panel show 
the numlocs=5 case, with results for both right- and left-tail dependence of the noise shown. 
And make the right panel show the numlocs=25 case with results for both right- and left-tail
dependence of the noise shown.
3) Right now you use grey dots on one panel and black on the other, but you can just use black 
on both.
4)  Use N instead of numlocs - in papers we use variable names like in math, not like in code.
***Shya: Done.-->

### 2019-02-18 SupportingInformation.Rmd
edited methods and background and some of the fig for extinction risk stuff
response added

```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}

@@

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.5),mgp=c(3,1,0))
s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
s1<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1.2,pch=1)
legend(x=0.2,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
       pch=c(16,1), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.2),
       horiz=F, bty='n')

plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1.2,pch=1,col="grey")
legend(x=0.2,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
       pch=c(16,1), lty=c(1,1), col=c("grey","grey"), cex=1.4,pt.cex=c(1.5,1.2),
       horiz=F, bty='n')

@@

# plot for global dispersal in D matrix 
pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.2),mgp=c(3,1,0))
s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
s1<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",xlab="d",ylab="Extinction risk",
     cex.lab=1.5,cex.axis=1.5,cex=1.5,pch=16,lwd=1.5)
lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1.2,pch=1)
legend(x=0.2,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
       pch=c(16,1), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.2),
       horiz=F, bty='n')

plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",xlab="d",ylab="Extinction risk",
     cex.lab=1.5,cex.axis=1.5,cex=1.5,pch=16,lwd=1.5)
lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1.2,pch=1,col="grey")
legend(x=0.2,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
       pch=c(16,1), lty=c(1,1), col=c("grey","grey"), cex=1.4,pt.cex=c(1.5,1.2),
       horiz=F, bty='n')

\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersion.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}

@@

@@

\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of dependence between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is an extremely commonly observed and widely studied 
[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means and variances are computed across space. Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015,@Reuman2017].

We examined using simulations whether copula structure influences spatial Taylor's law as 
follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
with unit diagonal entries 
and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
were organized into an $N \times n \times 1000$ array representing $1000$ simulations of 
populations measured in $n$ locations over $N$ time steps. 
The cumulative distribution function
(cdf) of the standard normal distribution was applied to these numbers, followed by the 
inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
These transformations made gamma distributed the marginal distributions of 
each population sampling location in each simulation. 
A skewed population marginal distribution is an 
important factor in Taylor's law [@Cohen2015,@Reuman2017]. We refer henceforth to the 
resulting $N \times n \times 1000$ array as our *N array* for this simulation exercise;
copula structure of dependences between populations in different sampling locations
was normal (N) for the simulations in this array.

@@

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{Mean curvature of the spatial Taylor's law relationship for 
corn leaf aphid abundance
data (A) was significantly less than for surrogates (B, C). The Taylor's law 
relationship for one example surrogate (B) shows shows greater curvature than
for data (A), and the significance of the difference was demonstrated
by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}
***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
***sHYA: dONE.
-->

>>>>

```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}

@@

sl5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
saveRDS(sl5,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")

sl25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
saveRDS(sl25,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")

sg5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
saveRDS(sg5,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")

sg25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
saveRDS(sg25,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
```

```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,sl5,sl25,sg5,sg25)}
sl5<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
sl25<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
sg5<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
sg25<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
plot(sl5$d_seq[1:6],sl5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sl5$d_seq[1:6],sl5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')

plot(sl25$d_seq[1:6],sl25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sl25$d_seq[1:6],sl25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')

@@

# plot for global dispersal in D matrix 
pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
plot(sg5$d_seq[1:6],sg5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sg5$d_seq[1:6],sg5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')

plot(sg25$d_seq[1:6],sg25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sg25$d_seq[1:6],sg25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')
par(op)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersal.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}

@@

# Extinction risk noise and dispersal matrix \label{ext_risk_model}

<!--How noise was generated ? code : ExtremeTailDep.R-->

<!--How dispersal matrix was generated ? code : mydispersal.R-->
A $N \times N$ dispersal matrix ($D$) was generated assuming that the $N$ number of metapopulation patches
were arranged in a linear chain fashion. Say, population from $i^{th}$ patch could either disperse to its nearest
connecting patches only (we called it as local dispersal) or could disperse to all other patches 
equally (we called it as global dispersal) at an dispersal rate, $d$. In case of 
local dispersal matrix the diagonal entries $D_{ii}$ would be equal to $d$ and its immediate neighbors
$D_{ii\pm j ,\;\ j=1}$ would get an equal proportion from the rest of $(1-d)$. The other distant neighbors 
$D_{ii\pm j ,\; j>1}$ were set to zero to execute local dispersal. For global dispersal matrix, $D_{ii}$
would be $d$ and all the other entries were equal to $(1-d)/N$. Code to generate these dispersal matrices 
was implemented in [Dispersal_mat](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R) function.

@@

\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of dependence between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is an extremely commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016] empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means and variances are computed across space. Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017].

We examined using simulations whether copula structure influences spatial Taylor's law as 
follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
with unit diagonal entries 
and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
were organized into an $N \times n \times 1000$ array representing $1000$ simulations of 
populations measured in $n$ locations over $N$ time steps. 
The cumulative distribution function
(cdf) of the standard normal distribution was applied to these numbers, followed by the 
inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
These transformations made gamma distributed the marginal distributions of 
each population sampling location in each simulation. 
A skewed population marginal distribution is an 
important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
resulting $N \times n \times 1000$ array as our *N array* for this simulation exercise;
copula structure of dependences between populations in different sampling locations
was normal (N) for the simulations in this array.

@@

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{Mean curvature of the spatial Taylor's law relationship for 
corn leaf aphid abundance
data (A) was significantly less than for surrogates (B, C). The Taylor's law 
relationship for one example surrogate (B) shows shows greater curvature than
for data (A), and the significance of the difference was demonstrated
by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}
***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
***Shya: dONE.
-->
